Developing effective theories for single and bilayer graphene using density matrix downfolding

### 2019-09-13 draft.tex
removed some unecessary acronyms

In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab initio} simulations has emerged.
Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation (RPA) or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
Instead, comparisons of model predictions to experiment are conducted for model validation.

@@

As such, a project named QMC High Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
My thesis will be focused on the first layer of downfolding.
I propose a two year plan for developing many-body lattice models for single- (SLG) and simple bi-layer (BLG) graphene systems using the DMD method and \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
My thesis will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and an eventual continuum model.

@@

The results of the linear regression are an estimate of the parameter values $g_k$ and model error $\epsilon[\Psi]$.
The error estimate usually takes the form of a goodness of fit, such as a cross-validated $R^2$ score or root mean squared error (RMSE).
These error estimates present a direct measurement of model quality, ensuring that DMD not only gives the best fit parameters but also an assessment of model validity.
Further, by quantifying model quality through statistical error estimates, it is possible to compare the accuracy of various low-energy \textit{ansatzes}.

@@

\begin{figure*}
\centering
%\hspace*{-0.25in}
\includegraphics[width=1.0\linewidth]{./figs/cuo_regr_2.pdf}
\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed parameters from the PCR, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
\label{fig:cuo_regr}
\end{figure*}

@@

\subsection{Constrained variational Monte Carlo (cVMC)}
A new constrained cVMC method will be used to sample the low-energy wave functions necessary for DMD.
The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
\begin{equation}
\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2
\end{equation}
with respect to the parameters $\vec{p}$.
The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^*$ is the target values of those descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}^*$.
This method allows one to target low-energy wave functions which sit near a target point $\vec{d}^*$ in descriptor space, allowing for a more systematic sampling of the low-energy space.

@@

\vspace{-10pt}
\paragraph{QMC HAMM}
A major goal of the QMC HAMM project is to develop a twist dependent, Moire length scale, interacting model for TBLG using QMC \textit{ab initio} calculations.
The first step in doing so is developing an atomic scale model for TBLG, and the necessary prerequisites for this step are a strong understanding of effective theories of the basic constituents of the system.
The models I propose to develop for SLG and AA/AB BLG provide exactly this kind of understanding.
As such, my work will lay the foundation for the four year QMC HAMM project to build an effective theory for TBLG.

>>>>

In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab initio} simulations has emerged.
Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
Instead, comparisons of model predictions to experiment are conducted for model validation.

@@

As such, a project named QMC High-Accuracy Multiscale Modeling (QMC HAMM) between research groups at UIUC, UCSB and Rice University is dedicated towards the development of an accurate continuum-scale interacting model for TBLG.
The model development consists of two layers of downfolding: first, developing an atomic scale lattice model using \textit{ab initio} calculations, and second, downfolding the resultant lattice model to a continuum length scale.
My thesis will be focused on the first layer of downfolding.
I propose a two year plan for developing many-body lattice models for single- (SLG) and simple bi-layer (BLG) graphene systems using the DMD method and \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
My thesis will lay the necessary ground work for the development of a fully twist angle dependent lattice model for TBLG, and an eventual continuum model.

@@

The results of the linear regression are an estimate of the parameter values $g_k$ and model error $\epsilon[\Psi]$.
The error estimate usually takes the form of a goodness of fit, such as a cross-validated $R^2$ score or root mean squared error.
These error estimates present a direct measurement of model quality, ensuring that DMD not only gives the best fit parameters but also an assessment of model validity.
Further, by quantifying model quality through statistical error estimates, it is possible to compare the accuracy of various low-energy \textit{ansatzes}.

@@

\begin{figure*}
\centering
%\hspace*{-0.25in}
\includegraphics[width=1.0\linewidth]{./figs/cuo_regr_2.pdf}
\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed model parameters, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
\label{fig:cuo_regr}
\end{figure*}

@@

\subsection{Constrained variational Monte Carlo (cVMC)}
A new constrained cVMC method will be used to sample the low-energy wave functions necessary for DMD.
The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$:
\begin{equation}
\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2
\end{equation}
with respect to the parameters $\vec{p}$.
The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^*$ is the target values of those descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}^*$.
This method allows one to target low-energy wave functions which sit near a target point $\vec{d}^*$ in descriptor space, providing a more systematic method for sampling the low-energy space.

@@

\vspace{-10pt}
\paragraph{QMC HAMM}
A major goal of the QMC HAMM project is to develop a twist dependent, Moire length scale, interacting model for TBLG using \textit{ab initio} QMC calculations.
The first step in doing so is developing an atomic scale model for TBLG, and the necessary prerequisites for this step are a strong understanding of effective theories of the basic constituents of the system.
The models I propose to develop for SLG and AA/AB BLG provide exactly this kind of understanding.
As such, my work will lay the foundation for the four year QMC HAMM project to build an effective theory for TBLG.
